---
title: "Modern Data Mining, HW 2"
author:
- Group Member 1
- Group Member 2
- Group Member 3
date: 'Due: 11:59 PM,  Sunday, 02/13'
output:
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: '4'
  html_document:
    code_folding: show
    highlight: haddock
    number_sections: yes
    theme: lumen
    toc: yes
    toc_depth: 4
    toc_float: yes
  word_document:
    toc: yes
    toc_depth: '4'
urlcolor: blue
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width=8, fig.height=4)
options(scipen = 0, digits = 3)  # controls base R output
# check if you have ISLR package, if not, install it
if(!require('pacman')) {install.packages('pacman')}
pacman::p_load(ISLR, tidyverse, data.table) # add the packages needed
```


\pagebreak

# Overview {-}

Principle Component Analysis is widely used in data exploration, dimension reduction, data visualization. The aim is to transform original data into uncorrelated linear combinations of the original data while keeping the information contained in the data. High dimensional data tends to show clusters in lower dimensional view. 

Clustering Analysis is another form of EDA. Here we are hoping to group data points which are close to each other within the groups and far away between different groups. Clustering using PC's can be effective. Clustering analysis can be very subjective in the way we need to summarize the properties within each group. 

Both PCA and Clustering Analysis are so called unsupervised learning. There is no response variables involved in the process. 

For supervised learning, we try to find out how does a set of predictors relate to some response variable of the interest. Multiple regression is still by far, one of the most popular methods. We use linear models as a working model for its simplicity and interpretability. It is important that we use domain knowledge as much as we can to determine the form of the response as well as the function format of the factors.


## Objectives

- PCA
- SVD
- Clustering Analysis
- Linear Regression

## Review materials

- Study Module 2: PCA
- Study Module 3: Clustering Analysis
- Study Module 4: Multiple regression

## Data needed

- `NLSY79.csv`
- `brca_subtype.csv`
- `brca_x_patient.csv`

# Case study 1: Self-seteem 

Self-esteem generally describes a person's overall sense of self-worthiness and personal value. It can play significant role in one's motivation and success throughout the life. Factors that influence self-esteem can be inner thinking, health condition, age, life experiences etc. We will try to identify possible factors in our data that are related to the level of self-esteem. 

In the well-cited National Longitudinal Study of Youth (NLSY79), it follows about 13,000 individuals and numerous individual-year information has been gathered through surveys. The survey data is open to public [here](https://www.nlsinfo.org/investigator/). Among many variables we assembled a subset of variables including personal demographic variables in different years, household environment in 79, ASVAB test Scores in 81 and Self-Esteem scores in 81 and 87 respectively. 

The data is store in `NLSY79.csv`.

Here are the description of variables:

**Personal Demographic Variables**

* Gender: a factor with levels "female" and "male"
* Education05: years of education completed by 2005
* HeightFeet05, HeightInch05: height measurement. For example, a person of 5'10 will be recorded as HeightFeet05=5, HeightInch05=10.
* Weight05: weight in lbs.
* Income87, Income05: total annual income from wages and salary in 2005. 
* Job87, Job05: job type in 1987 and 2005, including Protective Service Occupations, Food Preparation and Serving Related Occupations, Cleaning and Building Service Occupations, Entertainment Attendants and Related Workers, Funeral Related Occupations, Personal Care and Service Workers, Sales and Related Workers, Office and Administrative Support Workers, Farming, Fishing and Forestry Occupations, Construction Trade and Extraction Workers, Installation, Maintenance and Repairs Workers, Production and Operating Workers, Food Preparation Occupations, Setters, Operators and Tenders,  Transportation and Material Moving Workers
 
 
**Household Environment**
 
* Imagazine: a variable taking on the value 1 if anyone in the respondent’s household regularly read magazines in 1979, otherwise 0
* Inewspaper: a variable taking on the value 1 if anyone in the respondent’s household regularly read newspapers in 1979, otherwise 0
* Ilibrary: a variable taking on the value 1 if anyone in the respondent’s household had a library card in 1979, otherwise 0
* MotherEd: mother’s years of education
* FatherEd: father’s years of education
* FamilyIncome78

**Variables Related to ASVAB test Scores in 1981**

Test | Description
--------- | ------------------------------------------------------
AFQT | percentile score on the AFQT intelligence test in 1981 
Coding | score on the Coding Speed test in 1981
Auto | score on the Automotive and Shop test in 1981
Mechanic | score on the Mechanic test in 1981
Elec | score on the Electronics Information test in 1981
Science | score on the General Science test in 1981
Math | score on the Math test in 1981
Arith | score on the Arithmetic Reasoning test in 1981
Word | score on the Word Knowledge Test in 1981
Parag | score on the Paragraph Comprehension test in 1981
Numer | score on the Numerical Operations test in 1981

**Self-Esteem test 81 and 87**

We have two sets of self-esteem test, one in 1981 and the other in 1987. Each set has same 10 questions. 
They are labeled as `Esteem81` and `Esteem87` respectively followed by the question number.
For example, `Esteem81_1` is Esteem question 1 in 81.

The following 10 questions are answered as 1: strongly agree, 2: agree, 3: disagree, 4: strongly disagree

* Esteem 1: “I am a person of worth”
* Esteem 2: “I have a number of good qualities”
* Esteem 3: “I am inclined to feel like a failure”
* Esteem 4: “I do things as well as others”
* Esteem 5: “I do not have much to be proud of”
* Esteem 6: “I take a positive attitude towards myself and others”
* Esteem 7: “I am satisfied with myself”
* Esteem 8: “I wish I could have more respect for myself”
* Esteem 9: “I feel useless at times”
* Esteem 10: “I think I am no good at all”

## Data preparation

Load the data. Do a quick EDA to get familiar with the data set. Pay attention to the unit of each variable. Are there any missing values?


## Self esteem evaluation

Let concentrate on Esteem scores evaluated in 87. 

1. Reverse Esteem 1, 2, 4, 6, and 7 so that a higher score corresponds to higher self-esteem. (Hint: if we store the esteem data in `data.esteem`, then `data.esteem[,  c(1, 2, 4, 6, 7)]  <- 5 - data.esteem[,  c(1, 2, 4, 6, 7)]` to reverse the score.)

2. Write a brief summary with necessary plots about the 10 esteem measurements.

3. Do esteem scores all positively correlated? Report the pairwise correlation table and write a brief summary.

4. PCA on 10 esteem measurements. (centered but no scaling)

    a) Report the PC1 and PC2 loadings. Are they unit vectors? Are they orthogonal? 
  
    b) Are there good interpretations for PC1 and PC2? (If loadings are all negative, take the positive loadings for the ease of interpretation)
  
    c) How is the PC1 score obtained for each subject? Write down the formula.
    
    d) Are PC1 scores and PC2 scores in the data uncorrelated? 
    
    e) Plot PVE (Proportion of Variance Explained) and summarize the plot. 
  
    f) Also plot CPVE (Cumulative Proportion of Variance Explained). What proportion of the variance in the data is explained by the first two principal components?
  
    g) PC’s provide us with a low dimensional view of the self-esteem scores. Use a biplot with the first two PC's to display the data.  Give an interpretation of PC1 and PC2 from the plot. (try `ggbiplot` if you could, much prettier!)

5. Apply k-means to cluster subjects on the original esteem scores

    a) Find a reasonable number of clusters using within sum of squared with elbow rules.
    
    b) Can you summarize common features within each cluster?
    
    c) Can you visualize the clusters with somewhat clear boundaries? You may try different pairs of variables and different PC pairs of the esteem scores.


6. We now try to find out what factors are related to self-esteem? PC1 of all the Esteem scores is a good variable to summarize one's esteem scores. We take PC1 as our response variable. 

    a) Prepare possible factors/variables:

      - Personal information: gender, education (05, problematic), log(income) in 87, job type in 87, Body mass index as a measure of health (The BMI is defined as the body mass divided by the square of the body height, and is universally expressed in units of kg/m²). Since BMI is measured in 05, this will not be a good practice to be inclueded as possible variables. 
          
      - Household environment: Imagazine, Inewspaper, Ilibrary, MotherEd, FatherEd, FamilyIncome78. Do set indicators `Imagazine`, `Inewspaper` and `Ilibrary` as factors. 
    
      - Use PC1 of SVABS as level of intelligence
        
    b)   Run a few regression models between PC1 of all the esteem scores and suitable variables listed in a). Find a final best model with your own criterion. 

      - How did you land this model? Run a model diagnosis to see if the linear model assumptions are reasonably met. 
        
      - Write a summary of your findings. In particular, explain what and how the variables in the model affect one's self-esteem. 
        


# Case study 2: Breast cancer sub-type


[The Cancer Genome Atlas (TCGA)](https://www.cancer.gov/about-nci/organization/ccg/research/structural-genomics/tcga), a landmark cancer genomics program by National Cancer Institute (NCI), molecularly characterized over 20,000 primary cancer and matched normal samples spanning 33 cancer types. The genome data is open to public from the [Genomic Data Commons Data Portal (GDC)](https://portal.gdc.cancer.gov/).
 
In this study, we focus on 4 sub-types of breast cancer (BRCA): basal-like (basal), Luminal A-like (lumA), Luminal B-like (lumB), HER2-enriched. The sub-type is based on PAM50, a clinical-grade luminal-basal classifier. 

* Luminal A cancers are low-grade, tend to grow slowly and have the best prognosis.
* Luminal B cancers generally grow slightly faster than luminal A cancers and their prognosis is slightly worse.
* HER2-enriched cancers tend to grow faster than luminal cancers and can have a worse prognosis, but they are often successfully treated with targeted therapies aimed at the HER2 protein. 
* Basal-like breast cancers or triple negative breast cancers do not have the three receptors that the other sub-types have so have fewer treatment options.

We will try to use mRNA expression data alone without the labels to classify 4 sub-types. Classification without labels or prediction without outcomes is called unsupervised learning. We will use K-means and spectrum clustering to cluster the mRNA data and see whether the sub-type can be separated through mRNA data.

We first read the data using `data.table::fread()` which is a faster way to read in big data than `read.csv()`. 

```{r}
brca <- fread("data/brca_subtype.csv")

# get the sub-type information
brca_subtype <- brca$BRCA_Subtype_PAM50
brca <- brca[,-1]
```

1. Summary and transformation

    a) How many patients are there in each sub-type? 

    b) Randomly pick 5 genes and plot the histogram by each sub-type.

    c) Remove gene with zero count and no variability. Then apply logarithmic transform.


2. Apply kmeans on the transformed dataset with 4 centers and output the discrepancy table between the real sub-type `brca_subtype` and the cluster labels.

3. Spectrum clustering: to scale or not to scale?

    a) Apply PCA on the centered and scaled dataset. How many PCs should we use and why? You are encouraged to use `irlba::irlba()`.
    
    b) Plot PC1 vs PC2 of the centered and scaled data and PC1 vs PC2 of the centered but unscaled data side by side. Should we scale or not scale for clustering propose? Why? (Hint: to put plots side by side, use `gridExtra::grid.arrange()` or `ggpubr::ggrrange()` or `egg::ggrrange()` for ggplots; use `fig.show="hold"` as chunk option for base plots)

4. Spectrum clustering: center but do not scale the data

    a) Use the first 4 PCs of the centered and unscaled data and apply kmeans. Find a reasonable number of clusters using within sum of squared with the elbow rule.
    
    b) Choose an optimal cluster number and apply kmeans. Compare the real sub-type and the clustering label as follows: Plot scatter plot of PC1 vs PC2. Use point color to indicate the true cancer type and point shape to indicate the clustering label. Plot the kmeans centroids with black dots. Summarize how good is clustering results compared to the real sub-type.
    
    c) Compare the clustering result from applying kmeans to the original data and the clustering result from applying kmeans to 4 PCs. Does PCA help in kmeans clustering? What might be the reasons if PCA helps?
    
    d) Now we have an x patient with breast cancer but with unknown sub-type. We have this patient's mRNA sequencing data. Project this x patient to the space of PC1 and PC2. (Hint: remember we remove some gene with no counts or no variablity, take log and centered) Plot this patient in the plot in iv) with a black dot. Calculate the Euclidean distance between this patient and each of centroid of the cluster. Can you tell which sub-type this patient might have? 
    
```{r}
x_patient <- fread("data/brca_x_patient.csv")
```


# Case study 3: Auto data set

This question utilizes the `Auto` dataset from ISLR. The original dataset contains 408 observations about cars. It is similar to the CARS dataset that we use in our lectures. To get the data, first install the package ISLR. The `Auto` dataset should be loaded automatically. We'll use this dataset to practice the methods learn so far. 
Original data source is here: https://archive.ics.uci.edu/ml/datasets/auto+mpg

Get familiar with this dataset first. Tip: you can use the command `?ISLR::Auto` to view a description of the dataset. 

## EDA
Explore the data, with particular focus on pairwise plots and summary statistics. Briefly summarize your findings and any peculiarities in the data.
```{r}
library(ISLR)
?ISLR::Auto
pairs(Auto)
summary(Auto)
```
Japanese and European cars seem to be lighter and have lower horsepower,on average, when compared to their American counterparts. Displacement, horsepower, and weight all seem to be positively correlated. Cars with four cylinders seem to have higher MPG, on average, than cars with six and eight cylinders. Displacement, horsepower, and weight seem to be negatively correlated with MPG.

## What effect does `time` have on `MPG`?

a) Start with a simple regression of `mpg` vs. `year` and report R's `summary` output. Is `year` a significant variable at the .05 level? State what effect `year` has on `mpg`, if any, according to this model. 
```{r}
linreg1 <- lm(mpg ~ year, data = Auto)
summary(linreg1)
```
The hypothesis test which has H_0: year = 0 H_A: year != 0 has a p-value of approximately 0. We can confidently reject the null hypothesis and state that year is significant at the level of 0.05. The coefficient of year is 1.23 which means that every subsequent year a car is produced the MPG of cars produced that year increases by an average of 1.23.

b) Add `horsepower` on top of the variable `year` to your linear model. Is `year` still a significant variable at the .05 level? Give a precise interpretation of the `year`'s effect found here. 

```{r}
linreg2 <- lm(mpg ~ year + horsepower, data = Auto)
summary(linreg2)
```
The P-value of the hypothesis test is still under the level of 0.05 so we can confidently say that year is a significant variable. The coefficient has changed to 0.65727 which means that every subsequent year a car is produced the MPG of cars produced that year increases by an average of 0.65727.

c) The two 95% CI's for the coefficient of year differ among (i) and (ii). How would you explain the difference to a non-statistician?

```{r}
c(1.23 - 2*0.0874,1.23 + 2*0.0874)
c(0.65727 - 2*0.06626,0.65727 + 2*0.06626)
```
The 95% confidence interval for year according to the model created in part i is [1.06,1.40]. The meaning of this is that we are 95% confident in saying that every subsequent year a car is produced, the MPG of cars produced that year increases by a value in the range of [1.06,1.40]. The 95% confidence interval for year according to the model created in part ii is [0.525,0.790]. The meaning of this is that we are 95% confident in saying that every subsequent year a car is produced, the MPG of cars produced that year increases by a value in the range of [0.525,0.790].

d) Create a model with interaction by fitting `lm(mpg ~ year * horsepower)`. Is the interaction effect significant at .05 level? Explain the year effect (if any). 

```{r}
linreg3 <- lm(mpg ~ year * horsepower, data = Auto)
summary(linreg3)
```
The interaction effect is significant at the 0.05 level, the P-values of the F test is well below the significance threshold. When we multiply independent variables together in regression we lose our ability to make interpretations of coefficients.

## Categorical predictors

Remember that the same variable can play different roles! Take a quick look at the variable `cylinders`, and try to use this variable in the following analyses wisely. We all agree that a larger number of cylinders will lower mpg. However, we can interpret `cylinders` as either a continuous (numeric) variable or a categorical variable.

a) Fit a model that treats `cylinders` as a continuous/numeric variable. Is `cylinders` significant at the 0.01 level? What effect does `cylinders` play in this model?

```{r}
linreg4 <- lm(mpg ~ cylinders, data = Auto)
summary(linreg4)
```
The coefficient of cylinders is -3.558. This means that for every additional cylinder added to the engine block to car, on average, loses 3.558 units of MPG.

b) Fit a model that treats `cylinders` as a categorical/factor. Is `cylinders` significant at the .01 level? What is the effect of `cylinders` in this model? Describe the `cylinders` effect over `mpg`. 

```{r}
linreg5 <- lm(mpg ~ factor(cylinders), 
              data = Auto)
summary(linreg5)
```
The reference level for cylinders in this regression is 3. This means that all of the coefficients are compared to the base case, which is when there are three cylinders. The coefficient for 4 cylinders is 8.734. The interpretation of this is that compared to a car with 3 cylinders, cars with 4 cylinders have -- on average -- 8.734 more MPG. The same logic is extended to all cylinders in the dataset. Only the 4 cylinders inntercept is significant at the 0.01 significance level.

c) What are the fundamental differences between treating `cylinders` as a continuous and categorical variable in your models? 

The fundamental difference between treating a variable as continuous versus categorical is that coefficients for continuous variables are interpreted as describing the average change in the dependent variable when the independent variable is increased by one unit. Coefficients for categorical variables are interpreted as describing the average change when the variable is changed from the reference level to the level of the coefficient (changing from 3 to 4 cylinders in the example above).


d) Can you test the null hypothesis: fit0: `mpg` is linear in `cylinders` vs. fit1: `mpg` relates to `cylinders` as a categorical variable at .01 level?  

```{r}
fit0 <- lm(mpg ~ cylinders, data = Auto)
summary(fit0)


fit1 <- lm(mpg ~ factor(cylinders), data = Auto)
summary(fit1)


#F_Stat <- ((sum((fit0$residuals)^2) - sum((fit1$residuals)^2))/(390-387))/(sum((fit1$residuals)^2)/387)
#pf(F_Stat,3,387,lower.tail=F)
```

We cannot test that null hypothesis because regressing cylinders as a categorical variable against mpg is not a reduced form of the model where we are regressing cylinders as a continuous variable (and vice versa).

## Results

Final modeling question: we want to explore the effects of each feature as best as possible. You may explore interactions, feature transformations, higher order terms, or other strategies within reason. The model(s) should be as parsimonious (simple) as possible unless the gain in accuracy is significant from your point of view.
  
```{r}
finalModel <- lm(mpg ~ horsepower + weight + factor(cylinders) + year, data = Auto)
summary(finalModel)
plot(finalModel, 1)
plot(finalModel, 2)
```
  
a) Describe the final model. Include diagnostic plots with particular focus on the model residuals and diagnoses.
The model is homoskedastic and normally distributed, therefore we can trust its predictions. The only variables that were significant enough to keep in the model were weight, cylinders (as a categorical variable), and year (as a categorical variables). Adding other variables increased the R^2 slightly but it did not seem to be worth adding another parameter to the model.

b) Summarize the effects found.
The coefficient for weight is -0.006057, which means that each additional unit of weight added to the car decreases the mpg by an average of -0.006057. A car with 4 cylinders will have 7.161638 more mpg on average that a car with 3 cylinders. The coefficients for the rest of the number of cylinders in the dataset are as follows: 5 cylinders: 8.132421, 6 cylinders: 4.612818, 8 cylinders: 6.659609. The interpretations are the same as four cylinders but with their respective coefficients. Each subsequent year that a car is produced its mpg will increase by an average of 0.723.

c) Predict the `mpg` of the following car: A red car built in the US in 1983 that is 180 inches long, has eight cylinders, displaces 350 cu. inches, weighs 4000 pounds, and has a horsepower of 260. Also give a 95% CI for your prediction.
```{r}
newcar <- Auto[1,]
newcar[1:9] <- c(NA,8,350,260,4000,NA,83,1,NA)
pred <- predict(finalModel, newcar)
c(pred - 2*3.19,pred + 2*3.19) # 95% confidence interval
```

head(Auto)

# Simple Regression through simulations (option)
    
## Linear model through simulations

This exercise is designed to help you understand the linear model using simulations. In this exercise, we will generate $(x_i, y_i)$ pairs so that all linear model assumptions are met.

Presume that $\mathbf{x}$ and $\mathbf{y}$ are linearly related with a normal error $\boldsymbol{\varepsilon}$ , such that $\mathbf{y} = 1 + 1.2\mathbf{x} + \boldsymbol{\varepsilon}$. The standard deviation of the error $\varepsilon_i$ is $\sigma = 2$. 

We can create a sample input vector ($n = 40$) for $\mathbf{x}$ with the following code:

```{r, eval = F, echo = TRUE}
# Generates a vector of size 40 with equally spaced values between 0 and 1, inclusive
x <- seq(0, 1, length = 40)
```


### Generate data

Create a corresponding output vector for $\mathbf{y}$ according to the equation given above. Use `set.seed(1)`. Then, create a scatterplot with $(x_i, y_i)$ pairs. Base R plotting is acceptable, but if you can, please attempt to use `ggplot2` to create the plot. Make sure to have clear labels and sensible titles on your plots.


### Understand the model
i. Find the LS estimates of $\boldsymbol{\beta}_0$ and $\boldsymbol{\beta}_1$, using the `lm()` function. What are the true values of $\boldsymbol{\beta}_0$ and $\boldsymbol{\beta}_1$? Do the estimates look to be good? 

ii. What is your RSE for this linear model fit? Is it close to $\sigma = 2$? 

ii. What is the 95% confidence interval for $\boldsymbol{\beta}_1$? Does this confidence interval capture the true $\boldsymbol{\beta}_1$?

iii. Overlay the LS estimates and the true lines of the mean function onto a copy of the scatterplot you made above.


### diagnoses

i. Provide residual plot where fitted $\mathbf{y}$-values are on the x-axis and residuals are on the y-axis. 

ii. Provide a normal QQ plot of the residuals.

iii. Comment on how well the model assumptions are met for the sample you used. 



## Understand sampling distribution and confidence intervals

This part aims to help you understand the notion of sampling statistics and confidence intervals. Let's concentrate on estimating the slope only.  

Generate 100 samples of size $n = 40$, and estimate the slope coefficient from each sample. We include some sample code below, which should guide you in setting up the simulation. Note: this code is easier to follow but suboptimal; see the appendix for a more optimal R-like way to run this simulation.
```{r, eval = F, echo = TRUE}
# Inializing variables. Note b_1, upper_ci, lower_ci are vectors
x <- seq(0, 1, length = 40) 
n_sim <- 100              # number of simulations
b1 <- 0                   # n_sim many LS estimates of beta_1 (=1.2). Initialize to 0 for now
upper_ci <- 0             # upper bound for beta_1. Initialize to 0 for now.
lower_ci <- 0             # lower bound for beta_1. Initialize to 0 for now.
t_star <- qt(0.975, 38)   # Food for thought: why 38 instead of 40? What is t_star?

# Perform the simulation
for (i in 1:n_sim){I l
  y <- 1 + 1.2 * x + rnorm(40, sd = 2)
  lse <- lm(y ~ x)
  lse_output <- summary(lse)$coefficients
  se <- lse_output[2, 2]
  b1[i] <- lse_output[2, 1]
  upper_ci[i] <- b1[i] + t_star * se
  lower_ci[i] <- b1[i] - t_star * se
}
results <- as.data.frame(cbind(se, b1, upper_ci, lower_ci))

# remove unecessary variables from our workspace
rm(se, b1, upper_ci, lower_ci, x, n_sim, b1, t_star, lse, lse_out) 
```

i. Summarize the LS estimates of $\boldsymbol{\beta}_1$ (stored in `results$b1`). Does the sampling distribution agree with theory? 

ii.  How many of your 95% confidence intervals capture the true $\boldsymbol{\beta}_1$? Display your confidence intervals graphically. 




